---
title: A reprodicible report on Covid-19
output: html_document
---

```{r}
library(tidyverse) # metapackage of all tidyverse packages
library(ggplot2)
library(scales)
library(maps)
```

<font size="6"> Introduction </font>

The following analysis is conducted on the dataset of COVID-19 Data Repository by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University (JHU). In this study, the information of global confirmed cases and global death counts will be utilized. The dataset can be founded in https://github.com/CSSEGISandData/COVID-19

<font size="6"> Questions of Interest </font>

Q1: What are the top-10 countries with the most confirmed cases? <br>
Q2: What are the top-10 countries with the most death counts? <br>

```{r}
#options(warn=-1)
dir <- 'https://github.com/pjlau/Covid-19-Study/blob/main/'
confirmed_g <- 'RAW_global_confirmed_cases.csv'
death_g <- 'RAW_global_deaths.csv'
#gconf_dir <- paste(dir,confirmed_g,sep="")
#gconf <- read_csv(gconf_dir,show_col_types = FALSE,col_names = TRUE)

file_names <- c(confirmed_g,death_g)

urls <- str_c(dir, file_names)

gconf <- read_csv(urls[1])
gdeath <- read_csv(urls[2])
head(gconf,10)


#gdeath_dir <- paste(dir,death_g,sep="")
#gdeath <- read_csv(gdeath_dir,show_col_types = FALSE,col_names = TRUE)


```

```{r}
drop <- c("Province_State","Lat","Long")
gconf <- gconf[,!(names(gconf) %in% drop)]
head(gconf)
```

```{r}
dates <- names(gconf)[2:ncol(gconf)]
dates <- as.Date(dates,format='%m/%d/%y')
sort_idxs <- which(dates == "2022-12-31")

```

```{r}
drop <- c("Province_State","Lat","Long")
gdeath <- gdeath[,!(names(gdeath) %in% drop)]
head(gdeath,10)
```

<font size="6"> Bias identification: Records of some countries come with duplicate rows of entry. It is essential to keep the latest one with the most updated number. </font>

```{r}
aa <- gconf[order(gconf["Country_Region"], abs(gconf[ncol(gconf)]) ), ]
gconf <- gconf[!duplicated(aa["Country_Region"]), ]
dim(gconf)

bb <- gdeath[order(gdeath["Country_Region"], abs(gdeath[ncol(gdeath)]) ), ]
gdeath <- gdeath[!duplicated(bb["Country_Region"]), ]
dim(gdeath)

all.equal(gconf["Country_Region"], gdeath["Country_Region"])
```

```{r}
desc_gconf <- gconf[order(gconf[sort_idxs],decreasing=TRUE),]
head(desc_gconf,10)
top_10_conf <- desc_gconf[["Country_Region"]][1:10]
print(top_10_conf)
```

```{r}
desc_gdeath <- gdeath[order(gdeath[sort_idxs],decreasing=TRUE),]
head(desc_gdeath,10)
top_10_death <- desc_gdeath[["Country_Region"]][1:10]
print(top_10_death)
```

```{r}
dates <- names(gconf)[2:ncol(gconf)]
dates <- as.Date(dates,format='%m/%d/%y')
dates_range_idxs <- which(dates >= "2020-06-01" & dates <= "2022-12-31")
dates_range <- dates[dates_range_idxs]
dates_len <- length(dates)
dates_range_len <- length(dates_range_idxs)
#dates_range_idxs
countries <- gconf[["Country_Region"]]
countries_len <- length(countries)

top_conf_idxs <- c()
top_death_idxs <- c()
for (i in 1:10)
{
    top_conf_idxs <- c(top_conf_idxs,which(countries %in% top_10_conf[i]))
    top_death_idxs <- c(top_death_idxs,which(countries %in% top_10_death[i]))
}
#top_conf_idxs

#match(top_10_conf, top_conf_idxs)
#match(top_10_death, top_death_idxs)
```

```{r}
nconfd_cols <- c("Country_Region","Date","Confirmed","Death")
nconfd <- data.frame(matrix(nrow = dates_range_len*countries_len, ncol = length(nconfd_cols)))
colnames(nconfd) <- nconfd_cols

#dates_range

nconfd["Country/Region"]<- rep(countries, each = dates_range_len)
nconfd["Date"]<- rep(dates_range,times = countries_len)
nconfd["Confirmed"]<- as.vector(t(data.matrix(gconf[,dates_range_idxs])))
nconfd["Death"]<- as.vector(t(data.matrix(gdeath[,dates_range_idxs])))

head(nconfd)
dim(nconfd)
#nconfd[944,]
#nconfd[1888,]
#dim(nconfd_afg)
#ggplot(data=nconfd_afg, aes(x=Date, y=Confirmed,group = 1)) + geom_point() +
#  scale_x_date(date_breaks = "6 month", labels = date_format("%b %y"))
```

```{r}
nconf_top_cols <- append("Date",top_10_conf)
len_dates_range <- length(dates_range_idxs)
nconf_top <- data.frame(matrix(nrow = len_dates_range, ncol = length(nconf_top_cols)))
colnames(nconf_top) <- nconf_top_cols

date_len_idxs <- 1:length(dates_range_idxs)
#date_len_idxs
nconf_top[,1] <- dates_range
#length(dates_range_idxs)
for (i in 2:length(nconf_top_cols))
{
    curr_country <- top_conf_idxs[i-1]
    nconf_top[,nconf_top_cols[i]]  <-nconfd[(len_dates_range*(curr_country-1)+1):(len_dates_range*curr_country),3]
}
#top_conf_idxs
head(nconf_top)
```

<font size="6"> Figure 1: Top 10 countries of the confirmed cases as of 2022-12-31. </font>

```{r}
#ggplot(data=nconfd_afg, aes(x=Date, y=Confirmed,group = 1)) + geom_point() + scale_x_date(date_breaks = "6 month", labels = date_format("%b %y"))
nconf_top %>% 
  pivot_longer(!Date) %>%
  ggplot(aes(x = Date, y= value, color = name)) +
  geom_point()+ scale_x_date(date_breaks = "6 month", labels = date_format("%b %y"))
```

```{r}
ndeath_top_cols <- append("Date",top_10_death)
len_dates_range <- length(dates_range_idxs)
ndeath_top <- data.frame(matrix(nrow = len_dates_range, ncol = length(ndeath_top_cols)))
colnames(ndeath_top) <- ndeath_top_cols

date_len_idxs <- 1:length(dates_range_idxs)
#date_len_idxs
ndeath_top[,1] <- dates_range
#length(dates_range_idxs)
for (i in 2:length(ndeath_top_cols))
{
    curr_country <- top_death_idxs[i-1]
    ndeath_top[,ndeath_top_cols[i]]  <-nconfd[(len_dates_range*(curr_country-1)+1):(len_dates_range*curr_country),4]
}
#top_death_idxs
head(ndeath_top)
```

<font size="6"> Figure 2: Top 10 countries of the death counts as of 2022-12-31. </font>

```{r}
ndeath_top %>% 
  pivot_longer(!Date) %>%
  ggplot(aes(x = Date, y= value, color = name)) +
  geom_point()+ scale_x_date(date_breaks = "6 month", labels = date_format("%b %y"))
```

<font size="6">  Figure 3: Top 10 countries of the confirmed cases as of 2022-12-31 by a global heatmap. </font>

```{r}
map_top_cols <- top_10_conf
map_top_cols <- replace(map_top_cols, map_top_cols == 'US', 'USA')
map_top_cols <- replace(map_top_cols, map_top_cols == 'United Kingdom', 'UK')

lconf_top <- data.frame(
  map_id = map_top_cols,
  fill = as.matrix(unlist(nconf_top[len_dates_range,2:ncol(nconf_top)], use.names = FALSE))
)

#lconf_top

world_map <- map_data("world")
world_map <- subset(world_map, region != "Antarctica")

options(repr.plot.width = 10, repr.plot.height =6)

ggplot(lconf_top) +
  geom_map(
    dat = world_map, map = world_map, aes(map_id = region),
    fill = "white", color = "#7f7f7f", linewidth = 0.25
  ) +
  geom_map(map = world_map, aes(map_id = map_id, fill = fill), linewidth = 0.25) +
  scale_fill_gradient(low = "#fff7bc", high = "#cc4c02", name = "Total Confirmed Cases") +
  expand_limits(x = world_map$long, y = world_map$lat)+ggtitle("The Top Ten Countries As of 2022-12-31")
```

<font size="6"> Modeling: Predicting Death Counts from Confirmed Cases </font>

```{r}
# Select Rows by column value
nconfd_lastday <- nconfd[nconfd$Date == '2022-12-31',]
#nconfd_lastday

lm_lastday <- lm(Death~Confirmed, data = nconfd_lastday)
glm_lastday <- glm(Death~Confirmed, data = nconfd_lastday,family = "poisson")

# Adjusted R-squared value 
summary(lm_lastday)$adj.r.squared
# McFadden’s R-squared value
with(summary(glm_lastday), 1 - deviance/null.deviance)
```

R-squared values indicate that both the linear regression and poisson regression are generally good fits.

<font size = "6">Conclusion</font>

In this study, the top-10 countries with the most confirmed cases and death counts are identified. Line charts are plotted to show the total cases of those countries by date. A heatmap is applied to identify these countries geographically. Only 7 countries are among the top-10 of the most confirmed cases and death counts at the same time. Finally, regression models are created to assess if predicting death counts from confirmed cases could be a good fit. 

